Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law

### 2023-12-07 manuscript/manuscript.qmd
Add TNC and Greenpeace citations

---
title: "Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law"
short-title: "Operational constraints and closing civic space"
date: "2023-12-03"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  TODO: **Update this based on the finished version of the paper.** China’s 2016 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2016 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
word-count: 7938
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
toc-depth: 4
---

@@

INGOs with locally focused missions and connections with provincial officials are thus far more likely to purposely limit their activities to specific provinces, while organizations with broader missions have a larger geographic reach. While we cannot precisely tell which direction the causal story works with this phenomenon, we can make some reasonable speculations. It is likely that locally-oriented INGOs purposely network and seek out connections with local PSUs before formally registering, thus fostering better relationships with government officials and furthering the organization's mission within its limited provinces. It is also likely, given the high probability of national reach, that organizations without local connections purposely do not spend the time and effort to cultivate those relationships, as it is not necessary to operate across many provinces. A rival explanation that having local connections reduces an organization's operational space does not seem warranted—although INGOs with local connections do have a more limited geographic scope, organizations appear to purposely limit the number of provinces they work in in order to focus more on their local work.

@@

One key difference in these two organizations' differing registration strategies is their relationship with the government and alignment with government preferences. TNC has stated that its work is "science-based" and that the organization uses a "non-confrontational, collaborative approach in working closely with governments, the business community and others" [@ChinaPhilanthropyResearchInstitute:2018a]. Over its past two decades of work in China, TNC has provided technical support for the central government and local governments in Yunnan and Zhejiang Provinces, consulting on policies that have led to the establishment of national parks and nature reserves, and helping draft environmental policy proposals like the China National Biodiversity Conservation Strategy and Action Plan (2011–2030)[@todo]<https://www.nature.org/en-us/about-us/where-we-work/asia-pacific/china/stories-in-china/china-overview/>. This deliberately non-confrontational model, coupled with long-established partnerships with the government enabled TNC to secure the consent of a environment-focused national governmental agency to act as their PSU, which in turn allowed TNC to have a wider geographic reach throughout the country. Moreover, TNC's work with local and national governments has generally been aligned with government preferences and directly benefits the regime.

In contrast, Greenpeace engages in direct advocacy, "us[ing] peaceful protest and creative confrontation to expose global environmental problems" [@todo]<https://www.greenpeace.org/usa/about/>. The organization describes its work in China with phrases like "non-violent direct action," "impartiality and independence," and "brings change," indicating its dedication to maintaining organizational independence as it acts as an environmental watchdog [@todo]<https://www.greenpeace.org.cn/about/mission-and-value/>. In 2004, for instance, Greenpeace conducted three rounds of independent investigations into damages caused by the Sinar Mas Group, revealing how the Asian multinational corporation used promises of reforestation work as a cover while logging natural forests in Yunnan [@todo]<https://www.greenpeace.org.cn/2022/11/18/%e4%ba%a4%e9%94%8b%e4%ba%9a%e6%b4%b2%e6%9c%80%e5%a4%a7%e9%80%a0%e7%ba%b8%e5%b7%a8%e9%b3%84%ef%bc%8c%e5%ae%88%e6%8a%a4%e4%ba%91%e5%8d%97%e5%a4%a9%e7%84%b6%e6%9e%97-%e7%bb%bf%e8%89%b2%e5%92%8c/>. Greenpeace's ongoing independent advocacy work is often at odds with regime preferences, and from the government's perspective, its Chinese programming is unpredictable and less directly manageable. As a result, few potential PSUs have shown interest in sponsoring any potential registration attempt, and Greenpeace has chosen to work through repeated temporary activities rather than trying to gain RO status.

@@

We argue that regime preferences interact with organizations' issue areas to create variation in INGO operational space, or the *de facto* regulatory environment afforded to INGOs. We explore this theory by analyzing a novel dataset of all INGOs that have successfully registered in China since the implementation of the 2016 ONGO Law, a legal mechanism that provided the government with the ability to grow and shrink INGO operational space in strategic ways. By focusing on a specific law in one country, our work provides rich empirical illustrations of the mechanics of authoritarian administrative crackdown. We find that INGOs that work on issues that are less threatening and more aligned with regime preferences---such as arts & culture INGOs, industry associations, and scientific organizations—are registered in more provinces and thus enjoy a greater degree of operational space, while organizations that pose a potential threat---like education INGOs---are far more limited geographically. Our case study finds similar results. Because of its less confrontational approach and its connections to local government officials, The Nature Conservancy was able to obtain RO status and benefit from increased operational space while the more advocacy-focused Greenpeace has resorted to filing dozens of program-specific temporary activities.

>>>>

---
title: "Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law"
short-title: "Operational constraints and closing civic space"
date: "2023-12-07"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  TODO: **Update this based on the finished version of the paper.** China’s 2016 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2016 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
word-count: 7984
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
toc-depth: 4
---

@@

INGOs with locally focused missions and connections with provincial officials are thus far more likely to purposely limit their activities to specific provinces, while organizations with broader missions have a larger geographic reach. While we cannot precisely tell which direction the causal story works with this phenomenon, we can make some reasonable speculations. It is likely that locally-oriented INGOs purposely network and seek out connections with local PSUs before formally registering, thus fostering better relationships with government officials and furthering the organization's mission within its limited provinces. It is also likely, given the high probability of national reach, that organizations without local connections purposely do not spend the time and effort to cultivate those relationships, as it is not necessary to operate across many provinces. A rival explanation that having local connections reduces an organization's operational space does not seem warranted---although INGOs with local connections do have a more limited geographic scope, organizations appear to purposely limit the number of provinces they work in in order to focus more on their local work.

@@

One key difference in these two organizations' differing registration strategies is their relationship with the government and alignment with government preferences. TNC has stated that its work is "science-based" and that the organization uses a "non-confrontational, collaborative approach in working closely with governments, the business community and others" [@ChinaPhilanthropyResearchInstitute:2018a]. Over its past two decades of work in China, TNC has provided technical support for the central government and local governments in Yunnan and Zhejiang Provinces, consulting on policies that have led to the establishment of national parks and nature reserves, and helping draft environmental policy proposals like the China National Biodiversity Conservation Strategy and Action Plan (2011–2030) [@TheNatureConservancy:2023]. This deliberately non-confrontational model, coupled with long-established partnerships with the government enabled TNC to secure the consent of a environment-focused national governmental agency to act as their PSU, which in turn allowed TNC to have a wider geographic reach throughout the country. Moreover, TNC's work with local and national governments has generally been aligned with government preferences and directly benefits the regime.

In contrast, Greenpeace engages in direct advocacy, using peaceful---but confrontational---protests to expose environmental abuses and shame governments and for-profit corporations that harm the environment. The organization describes its work in China with phrases like "non-violent direct action," "impartiality and independence," and "brings change," indicating its dedication to maintaining organizational independence as it acts as an environmental watchdog [@Greenpeace:2023]. In 2004, for instance, Greenpeace conducted three rounds of independent investigations into damages caused by the Sinar Mas Group, revealing how the Asian multinational corporation used promises of reforestation work as a cover while logging natural forests in Yunnan [@Greenpeace:2022]. Greenpeace's ongoing independent advocacy work is often at odds with regime preferences, and from the government's perspective, its Chinese programming is unpredictable and less directly manageable. As a result, few potential PSUs have shown interest in sponsoring any potential registration attempt, and Greenpeace has chosen to work through repeated temporary activities rather than trying to gain RO status.

@@

We argue that regime preferences interact with organizations' issue areas to create variation in INGO operational space, or the *de facto* regulatory environment afforded to INGOs. We explore this theory by analyzing a novel dataset of all INGOs that have successfully registered in China since the implementation of the 2016 ONGO Law, a legal mechanism that provided the government with the ability to grow and shrink INGO operational space in strategic ways. By focusing on a specific law in one country, our work provides rich empirical illustrations of the mechanics of authoritarian administrative crackdown. We find that INGOs that work on issues that are less threatening and more aligned with regime preferences---such as arts & culture INGOs, industry associations, and scientific organizations---are registered in more provinces and thus enjoy a greater degree of operational space, while organizations that pose a potential threat---like education INGOs---are far more limited geographically. Our case study finds similar results. Because of its less confrontational approach and its connections to local government officials, The Nature Conservancy was able to obtain RO status and benefit from increased operational space while the more advocacy-focused Greenpeace has resorted to filing dozens of program-specific temporary activities.
